Perspective simulation

### 2021-02-11 scripts/r/simulations/simulations.Rmd
:sparkles: - use proper prior for sd

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(rmdformats)
library(tidyverse)
library(patchwork)
library(brms)
library(extraDistr)
library(betapart)
library(summarytools)
my_seed <- 115116106
```

# Meta-analysis

## Simulate data

@@

The $\eta_i$ is randomly sampled from a $Normal(1, 0.5)$ distribution, while $\text{se}_i$ from a $HalfCauchy(0, 0.1)$.

@@

# Number of teams, each team contributes one effect size
n <- 12
# standardized effect size
eta <- rnorm(n, 1, 0.5)
# standard error of the standardised effect size
se <- rhcauchy(n, 0.1)
# label each team
team <- letters[1:n]

@@

@@

<!-- TODO: need to discuss prior for sd -->

- A $Normal(0, 1)$ distribution for the prior of the intercept.
- A $HalfCauchy(0, 0.01)$ distribution for the prior of the intercept standard deviation.

```{r meta-priors}
priors <- c(
  prior(normal(0, 1), class = Intercept),
  prior(cauchy(0, 0.01), class = sd)
)
```

@@

```{r prior-predictive}
meta_bm_ppred <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  sample_prior = "only",
  chain = 1,
  seed = my_seed,
  file = "./models/sim/meta_bm_ppred"
)
```

```{r ppred-samples}
ppred_samples <- posterior_samples(meta_bm_ppred)
glimpse(ppred_samples)
```

@@

```{r meta-bm}
meta_bm <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  chain = 1,
  seed = my_seed,
  file = "./models/sim/meta_bm"
)
```

@@

# Simulated data summary

```{r dfsummary}
# seems dfSummary() does not support list columns
dfSummary(eta_tbl %>% select(-where(is.list)))
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(rmdformats)
library(tidyverse)
theme_set(theme_light())
library(patchwork)
library(brms)
library(extraDistr)
library(HDInterval)
library(betapart)
library(summarytools)
my_seed <- 115116106
```

# Simulate data

@@

The $\eta_i$ is randomly sampled from a $Normal(1, 0.5)$ distribution, while $\text{se}_i$ from a $HalfCauchy(0, 0.15)$.

@@

# Number of teams, each team contributes one effect size
n <- 12
# standardized effect size
eta <- rnorm(n, 1, 0.5)
# standard error of the standardised effect size
se <- rhcauchy(n, 0.15)
# label each team
team <- letters[1:n]

@@

# Meta-analytical model

## Model formula

$$
\begin{aligned}
\eta_i      & \sim \text{Normal}(\mu_i, \sigma_i) \\
\mu_i       & = \alpha + \alpha_{\text{t}[i]} \\
\alpha      & \sim \text{Normal}(0, 1) \\
\sigma_{\alpha_{\text{t}}} & \sim \text{HalfCauchy}(0, 1) \\
\sigma_i    & = \text{se}_i
\end{aligned}
$$

@@

- A $Normal(0, 1)$ distribution for the prior of the intercept.
- A $HalfCauchy(0, 1)$ distribution for the prior of the teams' intercept standard deviation.

```{r sd-prior}
tibble(
  x = seq(0, 10, by = 0.05),
  density = dhcauchy(x, 1)
) %>%
  ggplot(aes(x, density)) +
  geom_area(fill = "#325756", alpha = 0.9) +
  labs(
    title = "HalfCauchy(0, 1)",
    x = "teams' intercept standard deviation"
  )
```

```{r meta-priors}
priors <- c(
  prior(normal(0, 1), class = Intercept),
  prior(cauchy(0, 1), class = sd)
)
```

@@

```{r prior-predictive}
meta_bm_ppred <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  sample_prior = "only",
  chain = 4,
  seed = my_seed,
  file = "./models/sim/meta_bm_ppred"
)

ppred_samples <- posterior_samples(meta_bm_ppred) %>%
  rename(intercept = b_Intercept, `group-level SD` = sd_team__Intercept)
```

```{r ppred-posts}
ppred_samples %>%
  pivot_longer(intercept:`group-level SD`, names_to = "term") %>%
  mutate(term = factor(term, levels = c("intercept", "group-level SD"))) %>%
  ggplot(aes(value)) +
  geom_density(colour = NA, fill = "#325756", alpha = 0.9, bw = 0.4) +
  facet_wrap(~ term, scales = "free")
```

```{r ppred-joint-post}
ppred_samples %>%
  select(intercept, `r_team[a,Intercept]`:`r_team[l,Intercept]`) %>%
  pivot_longer(-intercept, names_to = "team") %>%
  mutate(eta = intercept + value) %>%
  ggplot(aes(eta)) +
  geom_density(colour = NA, fill = "#325756", alpha = 0.9, bw = 0.4) +
  xlim(-7, 7)
```

@@

```{r meta-bm}
meta_bm <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  chain = 4,
  seed = my_seed,
  file = "./models/sim/meta_bm"
)
```

@@
